\bf Eye Fitting Straight Lines in the Modern Era

### 2022-01-26 Eye-Fitting-Straight-Lines-in-the-Modern-Era.Rmd
added analysis code files to this repository.

eyefitting_example_sim <- read.csv(here("data/youdrawit-eyefitting-simdata-example.csv"))

@@

For each participant, the final data set used for analysis contains $x_{ijk}, y_{ijk,drawn}, \hat y_{ijk,OLS}$, and $\hat y_{ijk,PCA}$ for parameter choice $i = 1,2,3,4$, j = $1,...N_{participant}$, and $x_{ijk}$ value $k = 1, ...,4 x_{max} + 1$. 
Using both a linear mixed model and a generalized additive mixed model, comparisons of vertical residuals in relation to the OLS fitted values ($e_{ijk,OLS} = y_{ijk,drawn} - \hat y_{ijk,OLS}$) and PCA fitted values ($e_{ijk,PCA} = y_{ijk,drawn} - \hat y_{ijk,PCA}$) were made across the domain.
\cref{fig:eyefitting-example-plot} displays an example of all three fitted trend lines for parameter choice F. Data used in the analyses are available to be downloaded from GitHub [here](https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/raw/main/data/youdrawit-eyefitting-model-data.csv).

```{r eyefitting-example-plot, fig.height = 4, fig.width = 4, fig.cap = "Illustrates the data associated with and collected for one 'You Draw It' task plot. Trend-lines include the participant drawn line (dashed black), the OLS regression line (solid steelblue) and the PCA regression line based on the principal axis (solid orange).", out.width="50%"}
trial.feedback <- read.csv("data/youdrawit-eyefitting-example-feedback.csv") %>%
    mutate(`Parameter Choice` = "F")
trial.sim <- read.csv("data/youdrawit-eyefitting-example-simulated.csv") %>%
    mutate(`Parameter Choice` = "F")

@@

```{r eyefitting-lmer-residualplots, fig.height = 3, fig.width = 9, out.width = "100%", fig.cap = "Estimated trend line of the residuals between the participant drawn points and fitted values for both the OLS (blue) regression line and PCA (orange) regression line constrained to a linear fit modeled by a linear mixed model. Estimated residual trends with 95\\% confidence bands are overlaid on the observed individual participant residuals."}
eyefitting.preds.lmer <- read.csv("data/youdrawit-eyefitting-lmerpred-data.csv")
# Plot Predictions
eyefitting.lmer.plot <- eyefitting.preds.lmer %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualols, group = plotID, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualpca, group = plotID, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = asymp.LCL.ols, ymax = asymp.UCL.ols, fill = "OLS"), color = NA, alpha = 0.7) +
  geom_line(aes(y = emmean.ols, color = "OLS")) +
  geom_ribbon(aes(ymin = asymp.LCL.pca, ymax = asymp.UCL.pca, fill = "PCA"), color = NA, alpha = 0.7) +
  geom_line(aes(y = emmean.pca, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("LMER fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA"))

@@

```{r eyefitting-gamm-residualplots, fig.height = 3, fig.width = 9, out.width = "100%", fig.cap = "Estimated trend line of the residuals between the participant drawn points and fitted values for both the OLS (blue) regression line and PCA (orange) regression line determined by smoothing splines fit by a generalized additive mixed model. Estimated residual trends with 95\\% confidence bands are overlaid on the observed individual participant residuals."}
eyefitting.grid.gamm <- read.csv("data/youdrawit-eyefitting-gammpred-data.csv")
eyefitting.gamm.plot <- eyefitting.grid.gamm %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualols, group = plotID, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualpca, group = plotID, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = ols.lower, ymax = ols.upper, fill = "OLS"), color = NA, alpha = 0.5) +
  geom_line(aes(y = ols.pred, color = "OLS")) +
  geom_ribbon(aes(ymin = pca.lower, ymax = pca.upper, fill = "PCA"), color = NA, alpha = 0.5) +
  geom_line(aes(y = pca.pred, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("GAMM fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) 
eyefitting.gamm.plot
```

@@

\begin{description}

\item[Participant Data:] De-identified participant data can be found at ...

\item[Data Analysis Code:]

\end{description}

>>>>

eyefitting_example_sim <- read.csv(here("results/youdrawit-eyefitting-simdata-example.csv"))

@@

For each participant, the final data set used for analysis contains $x_{ijk}, y_{ijk,drawn}, \hat y_{ijk,OLS}$, and $\hat y_{ijk,PCA}$ for parameter choice $i = 1,2,3,4$, j = $1,...N_{participant}$, and $x_{ijk}$ value $k = 1, ...,4 x_{max} + 1$. 
Using both a linear mixed model and a generalized additive mixed model, comparisons of vertical residuals in relation to the OLS fitted values ($e_{ijk,OLS} = y_{ijk,drawn} - \hat y_{ijk,OLS}$) and PCA fitted values ($e_{ijk,PCA} = y_{ijk,drawn} - \hat y_{ijk,PCA}$) were made across the domain.
\cref{fig:eyefitting-example-plot} displays an example of all three fitted trend lines for parameter choice F.

```{r eyefitting-example-plot, fig.height = 4, fig.width = 4, fig.cap = "Illustrates the data associated with and collected for one 'You Draw It' task plot. Trend-lines include the participant drawn line (dashed black), the OLS regression line (solid steelblue) and the PCA regression line based on the principal axis (solid orange).", out.width="50%"}
trial.feedback <- read.csv("results/youdrawit-eyefitting-example-feedback.csv") %>%
    mutate(`Parameter Choice` = "F")
trial.sim <- read.csv("results/youdrawit-eyefitting-example-simulated.csv") %>%
    mutate(`Parameter Choice` = "F")

@@

```{r eyefitting-lmer-residualplots, fig.height = 3, fig.width = 9, out.width = "100%", fig.cap = "Estimated trend line of the residuals between the participant drawn points and fitted values for both the OLS (blue) regression line and PCA (orange) regression line constrained to a linear fit modeled by a linear mixed model. Estimated residual trends with 95\\% confidence bands are overlaid on the observed individual participant residuals."}
eyefitting.preds.lmer <- read.csv("results/youdrawit-eyefitting-lmerpred-data.csv")
# Plot Predictions
eyefitting.lmer.plot <- eyefitting.preds.lmer %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualols, group = plotID, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualpca, group = plotID, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = asymp.LCL.ols, ymax = asymp.UCL.ols, fill = "OLS"), color = NA, alpha = 0.7) +
  geom_line(aes(y = emmean.ols, color = "OLS")) +
  geom_ribbon(aes(ymin = asymp.LCL.pca, ymax = asymp.UCL.pca, fill = "PCA"), color = NA, alpha = 0.7) +
  geom_line(aes(y = emmean.pca, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("LMER fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA"))

@@

```{r eyefitting-gamm-residualplots, fig.height = 3, fig.width = 9, out.width = "100%", fig.cap = "Estimated trend line of the residuals between the participant drawn points and fitted values for both the OLS (blue) regression line and PCA (orange) regression line determined by smoothing splines fit by a generalized additive mixed model. Estimated residual trends with 95\\% confidence bands are overlaid on the observed individual participant residuals."}
eyefitting.grid.gamm <- read.csv("results/youdrawit-eyefitting-gammpred-data.csv")
eyefitting.gamm.plot <- eyefitting.grid.gamm %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualols, group = plotID, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data, aes(x = x, y = residualpca, group = plotID, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = ols.lower, ymax = ols.upper, fill = "OLS"), color = NA, alpha = 0.5) +
  geom_line(aes(y = ols.pred, color = "OLS")) +
  geom_ribbon(aes(ymin = pca.lower, ymax = pca.upper, fill = "PCA"), color = NA, alpha = 0.5) +
  geom_line(aes(y = pca.pred, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("GAMM fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) 
eyefitting.gamm.plot
```

@@

+ **Participant Data:** De-identified participant data collected in the study and used for analyses are available to be downloaded from GitHub [here](https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/tree/main/data). 
+ **Data Analysis Code:** The code used to replicate the analysis in this paper can be found [here](https://earobinson95.github.io/Eye-Fitting-Straight-Lines-in-the-Modern-Era/analysis/you-draw-it-eyefitting-analysis.html).
+ **Study Applet:** The shiny app used to conduct the study can be accessed [here](https://shiny.srvanderplas.com/you-draw-it/)

### 2022-01-26 Eye-Fitting-Straight-Lines-in-the-Modern-Era.tex
added analysis code files to this repository.

For each participant, the final data set used for analysis contains
\(x_{ijk}, y_{ijk,drawn}, \hat y_{ijk,OLS}\), and \(\hat y_{ijk,PCA}\)
for parameter choice \(i = 1,2,3,4\), j = \(1,...N_{participant}\), and
\(x_{ijk}\) value \(k = 1, ...,4 x_{max} + 1\). Using both a linear
mixed model and a generalized additive mixed model, comparisons of
vertical residuals in relation to the OLS fitted values
(\(e_{ijk,OLS} = y_{ijk,drawn} - \hat y_{ijk,OLS}\)) and PCA fitted
values (\(e_{ijk,PCA} = y_{ijk,drawn} - \hat y_{ijk,PCA}\)) were made
across the domain. \cref{fig:eyefitting-example-plot} displays an
example of all three fitted trend lines for parameter choice F. Data
used in the analyses are available to be downloaded from GitHub
\href{https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/raw/main/data/youdrawit-eyefitting-model-data.csv}{here}.

@@

\begin{description}

\item[Participant Data:] De-identified participant data can be found at ...

\item[Data Analysis Code:]

\end{description}

>>>>

For each participant, the final data set used for analysis contains
\(x_{ijk}, y_{ijk,drawn}, \hat y_{ijk,OLS}\), and \(\hat y_{ijk,PCA}\)
for parameter choice \(i = 1,2,3,4\), j = \(1,...N_{participant}\), and
\(x_{ijk}\) value \(k = 1, ...,4 x_{max} + 1\). Using both a linear
mixed model and a generalized additive mixed model, comparisons of
vertical residuals in relation to the OLS fitted values
(\(e_{ijk,OLS} = y_{ijk,drawn} - \hat y_{ijk,OLS}\)) and PCA fitted
values (\(e_{ijk,PCA} = y_{ijk,drawn} - \hat y_{ijk,PCA}\)) were made
across the domain. \cref{fig:eyefitting-example-plot} displays an
example of all three fitted trend lines for parameter choice F.

@@

\begin{itemize}
\tightlist
\item
  \textbf{Participant Data:} De-identified participant data collected in
  the study and used for analyses are available to be downloaded from
  GitHub
  \href{https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/tree/main/data}{here}.
\item
  \textbf{Data Analysis Code:} The code used to replicate the analysis
  in this paper can be found
  \href{https://earobinson95.github.io/Eye-Fitting-Straight-Lines-in-the-Modern-Era/analysis/you-draw-it-eyefitting-analysis.html}{here}.
\item
  \textbf{Study Applet:} The shiny app used to conduct the study can be
  accessed \href{https://shiny.srvanderplas.com/you-draw-it/}{here}
\end{itemize}
